---
title: "clean-rms"
author: "Suvit Verma"
date: "2026-05-27"
output: html_document
---

I am creating this document to clean and map RMS Installations with current SeM.

```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(purrr)
library(readxl)
library(stringr)
```

```{r}
df_sem<-read.csv("hfsem.csv")
df_sem<-df_sem|>select(-c(Open.Tickets,Closed.Tickets,No.of.Tickets.Logged))
df_sem<-df_sem|>mutate(norm_state=tolower(gsub(" ","",State)),norm_district=tolower(gsub(" ","",District)),norm_block=tolower(gsub(" ","",Block)))
```

```{r}
duplicates_all <- df |>group_by(Health.Facility.Name, NIN.OR.HFR, State, District, Block,Health.Facility.Type)|>filter(n() > 1)|>ungroup()
```
Duplicate -NIN or HFR
```{r}
duplicate_without_NIN<-df|>group_by(Health.Facility.Name, State, District, Block,Health.Facility.Type)|>filter(n() > 1)|>ungroup()
```

Duplicate District,State,District,Block
```{r}
duplicate_demo<- df |>group_by(Health.Facility.Name,State, District, Block)|>filter(n() > 1)|>ungroup()
```

See for duplicates in District Names

```{r}

duplicate_district <- df_sem %>%
  mutate(norm_district = tolower(gsub(" ", "", District))) %>%
  group_by(State, norm_district) %>%
  summarise(variations = list(unique(District)),
            count = n(),
            .groups = "drop") %>%
  filter(lengths(variations) > 1) %>%   # keep only groups with >1 spelling variation
  arrange(State, norm_district)
export_df <- duplicate_district %>%mutate(variations_str = sapply(variations, paste, collapse = ", ")) %>%
select(State, norm_district, variations_str)
write.csv(export_df, "duplicate_district.csv", row.names = FALSE)
```

Checking for Block
```{r}
df_sem <- df_sem %>%
  mutate(norm_state=tolower(gsub(" ","",State)),norm_district = tolower(gsub(" ", "", District)),norm_block=tolower(gsub(" ","",Block)),norm_type=tolower(gsub(" ","",Health.Facility.Type)),norm_name=tolower(gsub(" ","",Health.Facility.Name)))

duplicate_blocks <- df_sem %>%
  mutate(norm_block = tolower(gsub(" ", "", Block))) %>%
  group_by(State, norm_district, norm_block) %>%
  summarise(variations = list(unique(Block)),
            count = n(),
            .groups = "drop") %>%
  filter(lengths(variations) > 1) %>%
  arrange(State, norm_district, norm_block)
```

Checking Duplicacy in Health Center Type
```{r}
unique_hf_type<-unique(df_sem$Health.Facility.Type)
normalized_hf_type<-tolower(gsub(" ","",unique_hf_type))
dup_norm <- normalized_hf_type[duplicated(normalized_hf_type)]
dup_groups <- data.frame(original = unique_hf_type,
                         normalized_hf_type = normalized_hf_type) %>%
  group_by(normalized_hf_type) %>%
  summarise(variations = list(original), .groups = "drop") %>%
  filter(lengths(variations) > 1)
```

Here We will check if there are health center Names which repeat in different ways
```{r}
df_sem_non_alpha<-df_sem|>filter(grepl("[^A-Za-z ]",Health.Facility.Name))
df_sem_NIN_text<-df_sem|>filter(grepl("^[A-Za-z ]+$",NIN.OR.HFR))
df_sem_NIN_special<-df_sem|>filter(grepl("[^A-Za-z0-9 ]",NIN.OR.HFR))
df_sem_NIN_missing <- df_sem|>filter(NIN.OR.HFR == "(missing value)")
```


Correcting the & in Ayush Health & Wellness Center

```{r}
df_sem$Health.Facility.Type<-gsub("Ayush Health & Wellness Center","Ayush Health and Wellness Center",df_sem$Health.Facility.Type)
```

Cleaning
```{r}
# Dropping SC from norm_name wherever norm_type=SC
df_sem <- df_sem %>% mutate(norm_name = if_else(norm_type=="subcenter", str_replace(norm_name,"sc$",""), norm_name))

# Removing -Block Name from HF Name wherever Applicable
df_sem_block_name<-df_sem|>filter(str_detect(norm_name, paste0("-", norm_block, "$")))
df_sem <- df_sem %>%
  mutate(norm_name = if_else(str_detect(norm_name, paste0("-", norm_block, "$")),
                             str_replace(norm_name, paste0("-", norm_block, "$"), ""),
                             norm_name))
```



Now we will upload the RMS Installation Tracker

```{r}
df_rms<-read_excel("rms.xlsx",sheet = "tracker")

# Now in rms sheet I will replace Centre with Center

df_rms$`Type of health centre`<-gsub("Centre","Center",df_rms$`Type of health centre`)
df_rms$`Type of health centre`<-gsub("Health & Wellness Center","Health and Wellness Center",df_rms$`Type of health centre`)
df_rms<-df_rms|>select(`HFR ID/NIN`,`Name of health centre`,`State Name`,`District Name`,`Block Name`,`Type of health centre`,`Project Name/Code`,`Installation status`,`Date of Installation completed (DD-MM-YYYY)`,`RMS Meter Serial Number`,`RMS IP Number/MAC ID`)
```

Below is the list where there are more than one center
```{r}
duplicate_rms_all<-df_rms|>group_by(`Name of health centre`,`State Name`,`District Name`,`Block Name`,`Type of health centre`)|>filter(n()>1)|>ungroup()
```

Below are the center where in a block there are same center name
```{r}
duplicate_sdb_rms<-df_rms|>group_by(`Name of health centre`,`State Name`,`District Name`,`Block Name`)|>filter(n()>1)|>ungroup()
duplicate_sdb_sem<-df_sem|>group_by(Health.Facility.Name,State,District,Block)|>filter(n()>1)|>ungroup()
```

Checking duplicates district in the RMS sheet
```{r}
df_rms<-df_rms|>mutate(norm_state=tolower(gsub(" ","",`State Name`)),norm_district=tolower(gsub(" ","",`District Name`)),norm_block=tolower(gsub(" ","",`Block Name`)),norm_type=tolower(gsub(" ","",`Type of health centre`)),norm_name=tolower(gsub(" ","",`Name of health centre`)))

unique(df_rms$norm_state)

duplicate_rms_district<-df_rms|>group_by(norm_state,norm_district)|>summarise(variations = list(unique(`District Name`)),count = n(),.groups = "drop")|>filter(lengths(variations)>1)|>arrange(norm_state,norm_district)
```

Checking for duplicates in Block Name

```{r}
duplicate_rms_block<-df_rms|>group_by(norm_state,norm_district,norm_block)|>summarise(variations = list(unique(`Block Name`)),count = n(),.groups = "drop")|>filter(lengths(variations)>1)|>arrange(norm_state,norm_district,norm_block)
export_rms_block<-duplicate_rms_block|>mutate(variations_str = sapply(variations, paste, collapse = ", "))|>
select(norm_state,norm_district,norm_block, variations_str)
write.csv(export_rms_block,"duplicate_rms_block.csv",row.names = FALSE)
```

Cleaning Names in df_rms

```{r}

# Cleaning -block name
df_rms <- df_rms %>%
  mutate(norm_name = if_else(str_detect(norm_name, paste0("-", norm_block, "$")),
                             str_replace(norm_name, paste0("-", norm_block, "$"), ""),
                             norm_name))



df_rms <- df_rms %>%
  mutate(norm_name = str_trim(str_replace_all(norm_name, "subcentre", "")))
df_rms <- df_rms %>%
  mutate(norm_name = str_trim(str_replace(norm_name, "^phc", "")))
df_rms <- df_rms %>%
  mutate(norm_name = str_replace(norm_name, "primaryhealthcentre$", "phc"))
df_rms <- df_rms %>%
  mutate(norm_name = str_replace(norm_name, "subdivisionalhospital/th$", ""))
df_rms <- df_rms %>%
  mutate(norm_district = str_replace_all(norm_district, "ri-bhoi", "ribhoi"))
# Replacing the SC in df_rms
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="meghalaya" & norm_type %in% c("subcenter","healthandwellnesscenter"), str_replace(norm_name,"sc$",""), norm_name))

# In Meghalaya replacing subcenter with healthandwellness center

df_rms <- df_rms %>% mutate(norm_type = if_else(norm_state=="meghalaya" & norm_type=="subcenter","healthandwellnesscenter",norm_type))

# In Assam Cleaning for MPHC
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="assam" & norm_type=="miniprimaryhealthcenter", str_replace(norm_name,"mphc$",""), norm_name))
# Replacing all SC from end in subcenter norm_type in df_rms
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="subcenter", str_replace(norm_name,"sc$",""), norm_name))

#replacing SC with subcenter
df_rms <- df_rms %>% mutate(norm_type = if_else(norm_state=="arunachalpradesh" & norm_type=="sc","subcenter",norm_type))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="arunachalpradesh" & norm_type=="subcenter", str_replace(norm_name,"sc$",""), norm_name))

# Cleaning HWC in subcenter Arunachal
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="arunachalpradesh" & norm_type=="subcenter", str_replace(norm_name,"^hwc",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="arunachalpradesh" & norm_type=="subcenter", str_replace(norm_name,"hwc$",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="arunachalpradesh" & norm_type=="subcenter", str_replace(norm_name,"^sc",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="arunachalpradesh" & norm_type=="healthandwellnesscenter", str_replace(norm_name,"hwc$",""), norm_name))
#changing health and wellness center to subcenter
df_rms <- df_rms %>% mutate(norm_type = if_else(norm_state=="arunachalpradesh" & norm_type=="healthandwellnesscenter","subcenter",norm_type))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="arunachalpradesh" & norm_type=="primaryhealthcenter", str_replace(norm_name,"phc$",""), norm_name))

#replacing AHWC sites
df_rms <- df_rms %>% mutate(norm_type = if_else(norm_state=="arunachalpradesh" & str_detect(norm_name,"^ahwc"),"ayushhealthandwellnesscenter",norm_type))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="arunachalpradesh" & norm_type=="ayushhealthandwellnesscenter", str_replace(norm_name,"^ahwc",""), norm_name))

#replacing PHC in Primary Health Center
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"phc$",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"^phc",""), norm_name))

# For MPHC
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"^mphc",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"mphc$",""), norm_name))

#For SD
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"SD$",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"^SD",""), norm_name))

#For SHC
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"shc$",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="primaryhealthcenter", str_replace(norm_name,"^shc",""), norm_name))

# Cleaning HWC
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="assam" & norm_type=="healthandwellnesscenter", str_replace(norm_name,"hwc$",""), norm_name))
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_state=="assam" & norm_type=="healthandwellnesscenter", str_replace(norm_name,"^hwc",""), norm_name))

# cleaning clinic
df_rms <- df_rms %>% mutate(norm_name = if_else(norm_type=="healthclinic", str_replace(norm_name,"clinic$",""), norm_name))


```



```{r}
df_rms_mapped <- df_rms %>%
  left_join(
    df_sem %>%
      select(norm_state, norm_district, norm_block, norm_type, norm_name, NIN.OR.HFR),
    by = c("norm_state", "norm_district", "norm_block", "norm_type", "norm_name")
  ) %>%
  rename(NIN.HFR = NIN.OR.HFR)

df_unmapped <- df_rms_mapped %>%
  filter(is.na(NIN.HFR))

df_mapped<-df_rms_mapped|>filter(!is.na(NIN.HFR))
```

```{r}
df_temp_pending<-df_unmapped|>filter(is.na(`HFR ID/NIN`))
```

